---
title: "Writing Functions"
author: "Carolina Montes Garcia"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

library(tidyverse)
library(rvest)

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Here are some lists
```{r}
l = list(
  vec_numeric = 1:4,
  unif_sample = runif(100), 
  mat = matrix(1:8, nrow = 2, ncol = 4, byrow = TRUE), 
  summary = summary(rnorm(1000))
)

l

l$mat #just the matrix

l[["mat"]][1,3] #specific cell in the matrix

l[[1]]
l[[4]]
```

Make a list that's hopefully a bit more useful

```{r}
list_norm = 
  list(
    a = rnorm(20,0,5),
    b = rnorm(20,4,5),
    c = rnorm(20,0,10),
    d = rnorm(20,4,10)
  )

```

let's reuse the function
```{r}
mean_and_sd = function(x) {
  
  mean_x = mean(x)
  sd_x = sd(x)
  
  out_df = 
    tibble(
      mean = mean_x,
      sd = sd_x
    )

  return(out_df)
}
```

let's use the function to take mean and sd of all samples.

```{r}
mean_and_sd(list_norm[["a"]])
mean_and_sd(list_norm[["b"]])
mean_and_sd(list_norm[["c"]])
mean_and_sd(list_norm[["d"]])
```

## use a for loop
Create output list, and run a for loop

```{r}

output = vector("list", length = 4) #creating an empty list with 4 slots

for(i in 1:4){
  
  output[[i]] = mean_and_sd(list_norm[[i]])
  
}

output
```


## do the same thing 

but with `map` instead

```{r}

output = map(list_norm, mean_and_sd)

```

let's do a couple of other things

```{r}
output =
  map(list_norm, mean_and_sd)
bind_rows()

output = map_dfr(list_norm, mean_and_sd)
output = map_dbl(list_norm, IQR)
```
















## writing my first function!!
```{r}
x_vec = rnorm(n = 25, mean = 10, sd = 3.5)

(x_vec - mean(x_vec)) / sd(x_vec)
```

Now I'll write a function to do this

```{r}
z_scores = function(x){
  
  if(!is.numeric(x)){
    stop("x needs to be numeric")
  }
  
  if(length(x) <5){
    stop("you need at least 5 numbers to compute the z score")
  }
  
  z = (x - mean(x))/ sd(x)

  return(z)
  
}

z_scores(x = x_vec)
```

does this always work?

```{r, error = TRUE}
z_scores(x = 3)
z_scores(x = c("my", "name", "is", "carolina"))
```


```{r}
mean_and_sd = function(x) {
  
  mean_x = mean(x)
  sd_x = sd(x)
  
  out_df = 
    tibble(
      mean = mean_x,
      sd = sd_x
    )

  return(out_df)
}

mean_and_sd(x_vec)
```

## Check stuff using a simulation

```{r}
sim_df = 
  tibble(
    x = rnorm(30,10,5)
  )

sim_df %>% 
  summarize(
    mean= mean(x), 
    sd = sd(x)
  )
```

Simulation function to check sample mean and sd.

```{r}
#I can set default values for any arguments in my function and they can be overwritten later too. 
sim_mean_sd = function(samp_size, true_mean, true_sd) {
  
  sim_df = 
  tibble(
    x = rnorm(samp_size, true_mean, true_sd)
  )

out_df =
  sim_df %>% 
  summarize(
    mean= mean(x), 
    sd = sd(x)
  )
  
  return(out_df)
  
}

sim_mean_sd(samp_size = 30, true_mean = 4, true_sd = 12)
sim_mean_sd(samp_size = 30, true_mean = 4, true_sd = 12)

sim_mean_sd(30,16,2)

```

## Revisit LoTR words

```{r}

fellowship_df = 
  read_excel("data/LotR_words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship") %>% 
  janitor::clean_names()

two_towers_df = 
  read_excel("data/LotR_words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "two_towers")%>% 
  janitor::clean_names()

return_king_df = 
  read_excel("data/LotR_words.xlsx", range = "J3:L6") %>% 
  mutate(movie = "return_king")%>% 
  janitor::clean_names()
```

LEt's write a function for this process

```{r}
lotr_import = function(cell_range, movie_title) {
  
  movie_df = 
  read_excel("data/LotR_words.xlsx", range = cell_range) %>% 
  mutate(movie = movie_title) %>% 
  janitor::clean_names() %>% 
    pivot_longer(
      female:male,
      names_to = "sex",
      values_to = "words"
    )
  
  return(movie_df)
}

lotr_import("B3:B6", "fellowship")
lotr_import("F3:H6", "two_towers")
lotr_import("J3:L6", "return_king")


```

##NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

marj_table = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |>
  slice(-1) %>% 
  mutate(drug = "marj")


```

```{r}

nsduh_table = function(html, table_number, table_name){
  
  out_table = 
  html |>
  html_table() %>% 
  nth(table_number) |>
  slice(slide) %>% 
  mutate(table_name)
  
  return(out_table)
}

nsduh_table(html = nsduh_html, 1, "marj")
```

do this instead by saving function in an R script

```{r}
source("source/nsduh_table_format.R")

bind_rows(
  nsduh_table_format(html = nsduh_html, 1, "marj"),
  nsduh_table_format(html = nsduh_html, 4, "cocaine"),
  nsduh_table_format(html = nsduh_html, 5, "heroine")
)
```

